---
title: "Using SQL within R"
author: "Darrell Aucoin"
date: "February 13, 2015"
output: html_document
---

For SQL within R, we usually need two packages:

1. __DBI__: R Database Interface

2. Specific package for the individual SQL implementation (MySQL, SQLite, Oracle, etc.).  
  - Package `RSQLite` for SQLite

## DBI Functions

DBI contains various virtual classes and functions in connecting and querying a database:

__dbDriver__ Driver specifying the operations for creating connections to SQL Servers   

For SQLite:

```r
m = dbDriver("SQLite") # equivalent to SQLite()
```

For MySQL:  

```r
m = dbDriver("MySQL") # equivalent to MySQL()
```

__dbConnect__ Connect to a DBMS.  

- For SQLite:

```r
conn = dbConnect(m, "path_to_file")
```

- For MySQL

```r
conn = dbConnect(m, user="user_name", db="database_name", host="SQL_server", password="password")
```

__dbDisconnect__ Disconnect from a DBMS. You should always disconnect after you no longer need it

```r
dbDisconnect(conn)
```

__dbSendQuery__ Submits and executes SQL statement (information retrieved using fetch)  

```r
query = dbSendQuery(conn, "SELECT statement;")
```

__fetch__ Get records from a dbSendQuery

```r
res = fetch(query)
```

__dbClearResult__ _Free all resources_ (local and remote) from a query.

```r
dbClearResult(query)
```

__dbGetQuery__ Submits, executes SQL statement and retrieves information  

```r
res = dbGetQuery(conn, "SELECT statement;")
```

__dbListTables__ List tables in database connection

```r
tables = dbListTables(conn)
```

__dbGetInfo__ Get meta-data for DBIObjects

```r
meta.data = dbGetInfo(query)
```

__dbReadTable__ Fetch the data from a table.

```r
res = dbReadTable(conn, "table_name")
```

__dbListFields__ Return the column names for a given table

```
columns = dbListFields(conn, "table_name")
```

## Connecting a SQLite Database

```{r}
library(RSQLite)
library(xtable)
m = dbDriver("SQLite")
con = dbConnect(m, "/Users/darrellaucoin/GitHub/IntroSQL/stats_club.db")
query = dbSendQuery(con,
                   "SELECT *
                   FROM exec;")
result = fetch(query)
result
dbClearResult(query)
dbDisconnect(con)
```

---

```{r}
library(RSQLite)
m = dbDriver("SQLite")
con = dbConnect(m, "/Users/darrellaucoin/GitHub/IntroSQL/stats_club.db")
result = dbGetQuery(con,
                   "SELECT name, start_time, location
                   FROM event;")
result
dbDisconnect(con)
```

## Connecting to Local File System using MySQL

I transferred the data in stats_club.db to MySQL on my laptop.

```{r}
library("RMySQL") 
m = dbDriver("MySQL")
conn = dbConnect(m, user="darrell", db="stats_club",
                 host="localhost",
                 password="pass")
query = dbSendQuery(conn,
                    "SELECT name, start_time, location
                   FROM event")
result = fetch(query)
dbClearResult(query)
print(result)
dbDisconnect(conn)
```

## Connecting to Remote Server using MySQL

```{r}
hg19 = dbConnect(MySQL(), user="genome", db="hg19", 
	host="genome-mysql.cse.ucsc.edu")
result = dbGetQuery(hg19, 
		"SELECT COUNT(*) FROM affyU133Plus2")
dbDisconnect(hg19)
print(result)
```